---
title: "Compiled functions"
author: "zi jie"
date: "2022-11-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#install.packages("tidytext")
```


```{r}
library(dplyr)
library(tidyr)
library(tidyverse)
library(stringr)
library(knitr)
library(ggplot2)
library(ggmap)
library(leaflet)
library(leaflet.extras)
library(lubridate)
library(scales)
library(formattable)
library(rgdal)
library(raster)
library(shinyTime)
library(shiny)
library(shinycssloaders)
library(gganimate)
library(gifski)
library(png)
library(gapminder)
library(data.table)
library(rworldmap)
library(htmlwidgets)
library(XML)
library(jsonlite)
library(curl)
library(readxl)
library(mapview)
library(datefixR)
library(tm)
library(tidytext)
```


```{r}
df <- read.csv("finaldataset.csv", stringsAsFactors = FALSE)
zipcodes <- read.csv("sg_zipcode_mapper.csv")
```

#GENERALISED:
1. convert all blanks to NAs
```{r}
blank_NA <- function(df){
  df[df == ''] <- NA
  df
}
```

```{r}
df <- blank_NA(df)
```

2. fill postal 
```{r}
fill_postal <- function(df, colpostal, coladdress)
{
  zipcodes <- read.csv("sg_zipcode_mapper.csv")
  for (i in 1:nrow(df))
    {
    if (is.na(df[i, colpostal]))
    {
      filladdress <- df[i, coladdress]
      fillpostal <- zipcodes[which(zipcodes[, 8] == filladdress), 1]
      df[i, colpostal] <- fillpostal
    }
  }
  df[, colpostal] <- paste0("Singapore", " ", df[, colpostal])
  return(df)
}
```

```{r}
df <- fill_postal(df, 10,11)
```

3. fill address
```{r}
fill_address <- function(df, colpostal, coladdress)
{
  zipcodes <- read.csv("sg_zipcode_mapper.csv")
  for (i in 1:nrow(df))
    {
    if (is.na(df[i, coladdress]))
    {
      fillpostal <- str_extract(df[i, colpostal], "\\d+")
      filladdress <- zipcodes[which(zipcodes[, 1] == fillpostal), 8]
      df[i, coladdress] <- filladdress
    }
  }
  return(df)
}
```

```{r}
df <- fill_address(df, 10, 11)
```

4. add long/lat
```{r}
addLongLat <- function(df)
{
  df[, "Latitude"] <- NA
  df[, "Longitude"] <- NA
  for (i in 1:nrow(df))
    {
    df[i, "Latitude"] <- zipcodes[which(zipcodes[, 1] == str_extract(df[i, 10], "\\d+")), 2]
    df[i, "Longitude"] <- zipcodes[which(zipcodes[, 1] == str_extract(df[i, 10], "\\d+")), 3]
  }
  return(df)
}
```

```{r}
df <- addLongLat(df)
```

5. convert to lowercase
#changing all strings to small caps
```{r}
lower_string <- function(df){
  for(i in 1:ncol(df)) {# for-loop over columns
  if (class(df[, i]) == "character")
  df[ , i] <- tolower(df[ , i])
  }
  return(df)
}
```

```{r}
df <- lower_string(df)
```

6 & 7. replace then remove punc for cols except
```{r}
#for symptoms column
punctuation_replace <- function(df, colindex){
  library(stringr)
  library(stopwords)
  stopwords_regex = paste(stopwords("en"), collapse = "\\b|\\b")
  stopwords_regex = paste0("\\b", stopwords_regex, "\\b")
  df[, colindex] = stringr::str_replace_all(df[, colindex], stopwords_regex, ", ")
  a <- str_replace_all(df[, colindex], "[[:punct:]]", ", ")
}

#for all other columns
punctuation_remove <- function(df, colindex){
  df[, colindex] <- str_replace_all(df[, colindex], "[[:punct:]]", "")
}
```

```{r}
df$symptoms <- punctuation_replace(df, 12)
```

```{r}
df$case <- punctuation_remove(df, 2)
df$age <- punctuation_remove(df, 4)
df$gender <- punctuation_remove(df, 5)
df$nationality <- punctuation_remove(df, 6)
df$hospital <- punctuation_remove(df, 7)
df$postal <- punctuation_remove(df, 10)
```

8. convert age to numeric
```{r}
clean_age <- function(df, col1index){
  library(stringr) 
  #remove all punctuations
    df[,col1index] <- str_replace_all(df[,col1index], "[[:punct:]]", "") 
  #converting NAs to 0
  for (i in 1:nrow(df)){
      if (is.na(df[i,col1index])) {
    df[i,col1index] <- "0"
  }
  #cleaning dirty data
  df[,col1index] <- as.integer(df[,col1index])
    if (nchar(df[i,col1index]) >= 3 & abs(df[i,col1index]) > 113 & substr(df[i,col1index], 1, 1)!=0) {
    #converting back to integer format
df[i,col1index] <- as.integer(substr(df[i,col1index],1,2))
    }
  }
  return(df)
}
```

```{r}
df <- clean_age(df, 4)
```
9. fill in relevant NAs
```{r}
approx_NAs <- function(df, col1index)
{
  library(dplyr)
  discretetable <- data.frame()
  for (index in 1:length(table(df[,col1index])))
  {
    variable <- names(table(df[,col1index])[index])
    prob <- as.numeric(table(df[,col1index])[index])/length(df[,col1index])
    output <- c(variable, prob)
    discretetable <- rbind(discretetable, output)
  }
  colnames(discretetable) <- c("variable","prob")
  discretetable$prob <- as.numeric(discretetable$prob)
  
   dirty <- sample(x = discretetable$variable, 
                size = sum(is.na(df[,col1index])), replace = TRUE, 
                prob = discretetable$prob)

   df[,col1index] <- c(df[,col1index], dirty) %>% na.omit()
   warning("NAs have been replaced with values approximated by probability")
return(df)
}
```

```{r}
df <- approx_NAs(df,4)
df <- approx_NAs(df,5)
df <- approx_NAs(df,6)
df <- approx_NAs(df,7)
```
10. convert date
```{r}
library(devtools)
devtools::install_github("reconhub/linelist")
library(linelist)

std_date <- function(colname){
  a <- linelist::guess_dates(as.character(colname))
  a <- ymd(a)
}

df$date <- std_date(df$date)
```


```{r}
df$date <- std_date(df$date)
df$date.discharged <- std_date(df$date.discharged)
df$date.symptoms.onset <- std_date(df$date.symptoms.onset)
df$date.symptoms.stop <- std_date(df$date.symptoms.stop)
```

11. convert date.onset, date.stop
```{r}
validate_symptomsOnset <- function(df, colsymptom, colstart)
{
  for (i in 1:nrow(df)) {
    if (is.na(df[i, colstart]))
      if (!is.na(df[i, colsymptom]))
        df[i, colsymptom] <- NA
  }
  warning("Symptoms with no onset dates overwrite with NA")
  return(df)
}
```

```{r}
df <- validate_symptomsOnset(df,12,13)
```
12. validate stop dates
```{r}
validate_stopDates <- function(df, colstart, colstop)
{
  for (i in 1:nrow(df)) {
    if (is.na(df[i, colstart]))
    df[i, colstop] <- NA
  }
  warning("Date symptoms stop with no onset dates overwrite with NA")
  return(df)
}
```

```{r}
df <- validate_stopDates(df,13, 14)
```

13. typos overall
```{r}
mispell <- function(df){
  library(matchmaker)
  df <- df
  wordlist1 <- read.csv("spelling_dictionary.csv", 
                     stringsAsFactors = FALSE)
  df <- match_df(df, dictionary= wordlist1,
                                from = "options",
                                to = "values",
                                by = "group")
}
```
```{r}
df <- mispell(df)
```
```{r}
unique(df$nationality)
```


SPECIFIC:
14. gender typos
```{r}
clean_gender <- function(df, colgender)
{
   library(stringr)
  for (i in 1:nrow(df)){
     df[i,colgender] <- tolower(substr(df[i,colgender],1,1))
      if ((df[i,colgender])=="n") {
        df[i,colgender] <- "m"
        }
    }
 
  return(df)
  }
```

```{r}
df <- clean_gender(df, 5)
```


15. add is recovered col
```{r}
recovered <- function(df, coldischarged, coldeceased){
  df <- df
  df$is.recovered <- ifelse(!is.na(df[,coldischarged]) & df[,coldeceased] == FALSE| !is.na(df[,coldischarged]) & is.na(df[,coldeceased]), TRUE, FALSE)
  warning('Assumption that recovery is based on date discharge')
  return (df)
}
```

```{r}
df <- recovered(df, 8,9)
```


16. symptoms typo
```{r}
# Mispelt words for symptoms

clean_symptoms <- function(df, colsymptoms){
  df <- df
  numcol <- str_count(df[, colsymptoms], ", ") + 1
  max.col <- max(numcol, na.rm = T)
  split.col <- as.data.frame(trimws(str_split_fixed(df[, colsymptoms], pattern=", ", n=max.col)))
  df <- cbind(df, split.col)
  
  # Clean mispelt words
  library(matchmaker)
  wordlist1 <- read.csv("spelling_dictionary.csv", 
                     stringsAsFactors = FALSE)
  df <- match_df(df, dictionary= wordlist1,
                                from = "options",
                                to = "values",
                                by = "group")
  names <- colnames(split.col)
  df <- unite(df, col="symptoms", all_of(names), sep=", ", na.rm = TRUE)
  df
}
```

```{r}
df <- clean_symptoms(df, 12)
```


17. add symptoms duration
```{r}
addSymptomsDuration <- function(df, colstart, colstop, unit)
{
  df[, "Duration_Symptoms"] <- NA
  df[, "Duration_Symptoms"] <- difftime(df[, colstop], df[, colstart],  units = unit)
  if (TRUE %in% unique(df[, "Duration_Symptoms"] < 0))
    warning("Your symptoms duration data contains non-positive elements")
return(df)
}
```

```{r}
df <- addSymptomsDuration(df, 12, 13, "days")
```

18. add symptom count

```{r}
addsymptomCount <- function(df, colsymptom)
{
  df <- blank_NA(df)
  df[, "Symptoms_Count"] <- 0
  for (i in 1:nrow(df))
  {
    if (!is.na(df[i, colsymptom]) & str_count(df[i, colsymptom], ",") == 0)
    {df[i, "Symptoms_Count"] <- 1}
    else {df[i, "Symptoms_Count"] <- str_count(df[i, colsymptom], ",") + 1}
  }
  df[,"Symptoms_Count"] <- replace_na(df[,"Symptoms_Count"], 0)
  return(df)
}
```


```{r}
df <- addsymptomCount(df, 17)
```

```{r}
blank_NA <- function(df){
  df[df == ''] <- NA
  df
}
```

```{r}
df <- blank_NA(df)
```

19. count symptom type
```{r}
countSymptomType <- function(df, colsymptom)
{
  df[, colsymptom] <- gsub(" ", "", df[, colsymptom])
  symptoms_df <- data_frame(Text = df[, colsymptom])
  symptoms_words <- symptoms_df %>% unnest_tokens(output = word, 
                                                  input = Text) 
  symptoms_typecounts <- symptoms_words %>% 
    count(word, sort = TRUE)
  return(symptoms_typecounts)
}
```

```{r}
countSymptomType(df, 17)
```

```{r}
head(df[,-1],10)
```

```{r}
write.csv(df,"./cleansed_data.csv", row.names = FALSE)
```


END:
convert all col classes
